Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-01-27 reports/manuscript/paper.tex
Add some todo items, remove some todo items

\begin{abstract}
    % Abstracts should be unreferenced and typically around 150 words
    The white matter contains long-range connections between different
    brain regions and the organization of these connections holds important
    implications for brain function in health and disease. Tractometry
    uses diffusion-weighted magnetic resonance imaging (dMRI)
    to quantify tissue properties (e.g. fractional anisotropy (FA),
    mean diffusivity (MD), etc.), along the trajectories of these
    connections.
    % \cite{yeatman2012tract}.
    Statistical inference from
    tractometry usually either averages these quantities along
    the length of each bundle in each individual or performs analysis
    point-by-point along each bundle, with group comparisons or regression
    models computed separately for each point along every one of the bundles.
    These approaches are limited in their sensitivity, in the former case, or
    in their statistical power, in the latter.
    In the present work, we developed a method based on the sparse group
    lasso (SGL)
    % \cite{simon2013sparse}
    that takes into account tissue
    properties measured along all of the bundles, and selects informative
    features by enforcing sparsity, not only at the level of individual
    bundles, but also across the entire set of bundles and all of the measured
    tissue properties.
    % The sparsity penalties for each of these constraints
    % is identified using a nested cross-validation scheme that guards
    % against over-fitting and simultaneously identifies the correct
    % level of sparsity.
    We demonstrate the accuracy of the method in two
    settings: i) In a classification setting, patients with amyotrophic
    lateral sclerosis (ALS) are accurately distinguished from matched
    controls.
    % \cite{sarica2017corticospinal}.
    Furthermore, SGL automatically
    identifies FA in the corticospinal tract as important for this
    classification, correctly finding the parts of the white matter known
    to be affected by the disease. ii) In a regression setting, dMRI is
    used to accurately predict ``brain age.''
    % \cite{yeatman2014lifespan, Brown2012-so}.
    In this case, the weights are distributed throughout the
    white matter indicating that many different regions of the white matter
    change with development and contribute to the prediction of age. Thus,
    SGL makes it possible to leverage the multivariate relationship between
    diffusion properties measured along multiple bundles to make accurate
    predictions of subject characteristics while simultaneously discovering
    the most relevant features of the white matter for the characteristic of
    interest.
\end{abstract}

@@

\todo[inline]{The introduction section should be shortened for Nature methods
I think.}

@@

Methods and associated references are available in the supplemental material
\cite{supplement}.

>>>>

\begin{abstract}
    % Abstracts should be unreferenced and typically around 150 words
    \todo{Shorten abstract to roughly 150 words.}
    The white matter contains long-range connections between different
    brain regions and the organization of these connections holds important
    implications for brain function in health and disease. Tractometry
    uses diffusion-weighted magnetic resonance imaging (dMRI)
    to quantify tissue properties (e.g. fractional anisotropy (FA),
    mean diffusivity (MD), etc.), along the trajectories of these
    connections.
    % \cite{yeatman2012tract}.
    Statistical inference from
    tractometry usually either averages these quantities along
    the length of each bundle in each individual or performs analysis
    point-by-point along each bundle, with group comparisons or regression
    models computed separately for each point along every one of the bundles.
    These approaches are limited in their sensitivity, in the former case, or
    in their statistical power, in the latter.
    In the present work, we developed a method based on the sparse group
    lasso (SGL)
    % \cite{simon2013sparse}
    that takes into account tissue
    properties measured along all of the bundles, and selects informative
    features by enforcing sparsity, not only at the level of individual
    bundles, but also across the entire set of bundles and all of the measured
    tissue properties.
    % The sparsity penalties for each of these constraints
    % is identified using a nested cross-validation scheme that guards
    % against over-fitting and simultaneously identifies the correct
    % level of sparsity.
    We demonstrate the accuracy of the method in two
    settings: i) In a classification setting, patients with amyotrophic
    lateral sclerosis (ALS) are accurately distinguished from matched
    controls.
    % \cite{sarica2017corticospinal}.
    Furthermore, SGL automatically
    identifies FA in the corticospinal tract as important for this
    classification, correctly finding the parts of the white matter known
    to be affected by the disease. ii) In a regression setting, dMRI is
    used to accurately predict ``brain age.''
    % \cite{yeatman2014lifespan, Brown2012-so}.
    In this case, the weights are distributed throughout the
    white matter indicating that many different regions of the white matter
    change with development and contribute to the prediction of age. Thus,
    SGL makes it possible to leverage the multivariate relationship between
    diffusion properties measured along multiple bundles to make accurate
    predictions of subject characteristics while simultaneously discovering
    the most relevant features of the white matter for the characteristic of
    interest.
\end{abstract}

@@

@@

Methods and associated references are available in the supplemental material.

### 2021-01-27 reports/manuscript/supplemental.tex
Add some todo items, remove some todo items

Here, we propose that in addition to regularizing the coefficients in
$\hat{\beta}$, we can also capitalize on our knowledge of the group structure
of the bundle profile features in $\mathbf{X}$. The bundle-metric
combinations form a natural grouping. For example, we expect that MD features
within the left arcuate fasciculus will co-vary across individuals. Likewise
for FA values within the right corticospinal tract (CST) and so on. This
group structure is represented in \cref{main:fig:methods:group-structure}, which
depicts the linear model $\hat{y} = \mathbf{X} \hat{\beta}$. Thus, we seek a
regularization approach that will fit a linear model with
anatomically-grouped covariates, where we expect to observe both groupwise
sparsity, where the number of groups (bundle/metric combinations) with at
least one non-zero coefficients is small, as well as within-group sparsity,
where the number of non-zero coefficients within each non-zero group is
small. The sparse group lasso (SGL) is a penalized regression technique that
satisfies these criteria\cite{simon2013sparse}. It solves for a
coefficient vector $\hat{\beta}$ that satisfies
\begin{multline}
    \hat{\beta} = \min_\beta L_{\text{mse}}
    + \left( 1 - \alpha \right) \lambda \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} \norm{\beta^{(\ell)}}_2
    + \alpha \lambda \norm{\beta}_1,\\%
    \text{where} \quad
    L_{\text{mse}} = \frac{1}{2}
    \norm{y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \beta^{(\ell)}}_2^2,
    \label{eq:sgl}
\end{multline}
$G$ is the number of groups $\mathbf{X}^{(\ell)}$ is the submatrix of
$\mathbf{X}$ corresponding to group $\ell$, $\beta^{(\ell)}$ is the
coefficient vector for group $\ell$ and $p_\ell$ is the length of
$\beta^{(\ell)}$. In the tractomtetry setting, $G = T \times M$ and $\forall
\ell: p_\ell = 100$. The first term is the mean square error loss,
$L_{\text{mse}}$, as in the standard linear regression framework. The second
and third terms encourage groupwise sparsity and overall sparsity,
respectively. If $\alpha = 1$ the SGL reduces to the traditional
lasso\cite{tibshirani1996regression}. Conversely, if $\alpha = 0$ the SGL
reduces to the group lasso\cite{yuan2006model}. Thus, the model
hyperparameter $\alpha$ controls the combination of group-lasso and lasso.
The hyperparameter $\lambda$ controls the strength of the regularization.

@@

\Cref{fig:cc-bp:fa,fig:cc-bp:md} show the bundle profiles and regression
coefficients for the Cam-CAN dataset. Like the WH and HBN datasets, the
$\hat{\beta}$ coefficients are distributed widely through the brain and SGL
behaves more like the lasso than the group lasso. As before, one must be
cautious about comparing bundle profiles and $\hat{\beta}$ coefficients
between models. While the HBN and Cam-CAN datasets share the same diffusion
model and refrain from clipping streamlines, the age distributions for the
two are roughly disjoint, with the WH age distribution straddling the two.
The parsimony pitfall is again evident in the UNCL, UNCR, ARCL, SLFL, and
SLFR bundles in \cref{fig:cc-bp:md}.
\todo[inline]{We could say more here but it's not an ``aging'' paper so I
don't really want to. I only want to stress that you have to be careful
comparing $\hat{\beta}$ across models. Ariel: agreed!}

>>>>

Here, we propose that in addition to regularizing the coefficients in
$\hat{\beta}$, we can also capitalize on our knowledge of the group structure
of the bundle profile features in $\mathbf{X}$. The bundle-metric
combinations form a natural grouping. For example, we expect that MD features
within the left arcuate fasciculus will co-vary across individuals. Likewise
for FA values within the right corticospinal tract (CST) and so on. This
group structure is represented in \cref{main:fig:methods:group-structure}, which
depicts the linear model $\hat{y} = \mathbf{X} \hat{\beta}$. Thus, we seek a
regularization approach that will fit a linear model with
anatomically-grouped covariates, where we expect to observe both groupwise
sparsity, where the number of groups (bundle/metric combinations) with at
least one non-zero coefficients is small, as well as within-group sparsity,
where the number of non-zero coefficients within each non-zero group is
small. The sparse group lasso (SGL) is a penalized regression technique that
satisfies these criteria\cite{simon2013sparse}. It solves for a
coefficient vector $\hat{\beta}$ that satisfies
\begin{multline}
    \hat{\beta} = \min_\beta L_{\text{mse}}
    + \left( 1 - \alpha \right) \lambda \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} \norm{\beta^{(\ell)}}_2
    + \alpha \lambda \norm{\beta}_1,\\%
    \text{where} \quad
    L_{\text{mse}} = \frac{1}{2}
    \norm{y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \beta^{(\ell)}}_2^2.
    \label{eq:sgl}
\end{multline}
Here, $G$ is the number of groups, $\mathbf{X}^{(\ell)}$ is the submatrix of
$\mathbf{X}$ corresponding to group $\ell$, $\beta^{(\ell)}$ is the
coefficient vector for group $\ell$ and $p_\ell$ is the length of
$\beta^{(\ell)}$. In the tractomtetry setting, $G = T \times M$ and $\forall
\ell: p_\ell = 100$. The first term is the mean square error loss,
$L_{\text{mse}}$, as in the standard linear regression framework. The second
and third terms encourage groupwise sparsity and overall sparsity,
respectively. If $\alpha = 1$ the SGL reduces to the traditional
lasso\cite{tibshirani1996regression}. Conversely, if $\alpha = 0$ the SGL
reduces to the group lasso\cite{yuan2006model}. Thus, the model
hyperparameter $\alpha$ controls the combination of group-lasso and lasso.
The hyperparameter $\lambda$ controls the strength of the regularization.

@@

\Cref{fig:cc-bp:fa,fig:cc-bp:md} show the bundle profiles and regression
coefficients for the Cam-CAN dataset. Like the WH and HBN datasets, the
$\hat{\beta}$ coefficients are distributed widely through the brain and SGL
behaves more like the lasso than the group lasso. As before, one must be
cautious about comparing bundle profiles and $\hat{\beta}$ coefficients
between models. While the HBN and Cam-CAN datasets share the same diffusion
model and refrain from clipping streamlines, the age distributions for the
two are roughly disjoint, with the WH age distribution straddling the two.
The parsimony pitfall is again evident in the UNCL, UNCR, ARCL, SLFL, and
SLFR bundles in \cref{fig:cc-bp:md}.
